---
title: "Voting Machine Breakdown"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(wesanderson)


options(scipen=999)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
techlog <- read.csv('Machine Technician Log.csv')
vollog <-read.csv('Machine Volunteer Log.csv')
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
techlog$WARD_DIV <- paste(techlog$WARD,techlog$DIVISION)
techlog$WARD <- as.character(techlog$WARD)
techlog$DIVISION <- as.character(techlog$DIVISION)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height = 8, fig.width = 10}
ggplot(aes(x=reorder(WARD, -table(WARD)[WARD])), data=techlog) +
  geom_bar(fill=wes_palette(n=1, name="Chevalier")) +
  scale_y_continuous(breaks=c(1,2,3,4,5,6,7,8,9)) +
  labs(x='Ward Number', y='Number of Breakdowns')

```

Number of breakdowns by Ward, ordered highest to lowest. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
length(unique(techlog[["WARD_DIV"]]))
```

192 Divisions (out of 1686) were ultimately affected, going by the repair log

```{r echo=FALSE, message=FALSE, warning=FALSE}
vollog.clean <- subset(vollog, select = c('id', 'incident_time', 'scope', 'precinct'))
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=reorder(scope, -table(scope)[scope])), data=vollog.clean) +
  geom_bar(fill=wes_palette(n=1, name="Chevalier")) +
  labs(x='Scope of Affected Voters', y='Number of Breakdowns')
```

Most of the outages with an explicit scope affected more than one voter 
(going by volunteer log).

Not all breakdowns have a listed scope, though guesses could be made by checking
 notes, etc.

```{r echo=FALSE, message=FALSE, warning=FALSE}
vollog.clean$incident_time = gsub('2016-11-08',"",vollog.clean$incident_time)
vollog.clean$incident_time = gsub('EST',"",vollog.clean$incident_time)

```


```{r eval=FALSE, include=FALSE, include=FALSE}
ggplot(vollog.clean, aes(x = incident_time, y=time)) + 
  geom_line()
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height = 10, fig.width = 11}
ggplot(aes(x=reorder(COMPLAINT, -table(COMPLAINT)[COMPLAINT])), data=techlog) +
  geom_bar(fill=wes_palette(n=1, name="Chevalier")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x='Complaint', y='Number of Breakdowns')

```

Error Code 70 and Power Failure were the cause of the most breakdowns by far.


```{r}
#create dataset for most-affected wards (top ten?) and examine breakdowns
```


